---
title: "Analysis of Crime within the District of Columbia"
author: "Ashley Frankenfield, Gessica Vasconcelos"
date: "4/3/2020"
output: 
  rmdformats::readthedown:
    html_document:
      number_sections: TRUE
      self_contained: TRUE
      lightbox: TRUE
      gallery: FALSE
      highlight: tango 
      thumbnails: true
      toc_float: TRUE
      use_bookdown: TRUE
      code_folding: hide 
      max-width: 1800px
    pdf_document:
      toc: yes
      toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'hide', message = F, echo=F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
```

```{r basicfcn}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# note that using this function requires quotes around the package name, as you would when installing packages.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
# unload/detact package when done using it 
detach_package = function(pkg, character.only = FALSE) { if(!character.only) { pkg <- deparse(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```


```{r outlierKD2}
outlierKD2 <- function(df, var, rm=FALSE) { 
    #' Original outlierKD functino by By Klodian Dhana,
    #' https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/
    #' Modified to have third argument for removing outliers inwtead of interactive prompt, 
    #' and after removing outlier, original df will not be changed. The function returns the new df, 
    #' which can be saved as original df name if desired.
    #' Check outliers, and option to remove them, save as a new dataframe. 
    #' @param df The dataframe.
    #' @param var The variable in the dataframe to be checked for outliers
    #' @param rm Boolean. Whether to remove outliers or not.
    #' @return The dataframe with outliers replaced by NA if rm==TRUE, or df if nothing changed
    #' @examples
    #' outlierKD2(mydf, height, FALSE)
    #' mydf = outlierKD2(mydf, height, TRUE)
    #' mydfnew = outlierKD2(mydf, height, TRUE)
    dt = df # duplicate the dataframe for potential alteration
    var_name <- eval(substitute(var),eval(dt))
    na1 <- sum(is.na(var_name))
    m1 <- mean(var_name, na.rm = T)
    par(mfrow=c(2, 2), oma=c(0,0,3,0))
    boxplot(var_name, main="With outliers")
    hist(var_name, main="With outliers", xlab=NA, ylab=NA)
    outlier <- boxplot.stats(var_name)$out
    mo <- mean(outlier)
    var_name <- ifelse(var_name %in% outlier, NA, var_name)
    boxplot(var_name, main="Without outliers")
    hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
    title("Outlier Check", outer=TRUE)
    na2 <- sum(is.na(var_name))
    cat("Outliers identified:", na2 - na1, "\n")
    cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "\n")
    cat("Mean of the outliers:", round(mo, 2), "\n")
    m2 <- mean(var_name, na.rm = T)
    cat("Mean without removing outliers:", round(m1, 2), "\n")
    cat("Mean if we remove outliers:", round(m2, 2), "\n")
    
    # response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
    # if(response == "y" | response == "yes"){
    if(rm){
        dt[as.character(substitute(var))] <- invisible(var_name)
        #assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
        cat("Outliers successfully removed", "\n")
        return(invisible(dt))
    } else {
        cat("Nothing changed", "\n")
        return(invisible(df))
    }
}
```

```{r loadPkg}
  loadPkg("corrplot")
  loadPkg("dplyr")
  loadPkg("ggplot2")
  loadPkg("tidyverse")
  loadPkg("sqldf")
  loadPkg('tidyr')
```

## Introduction 
``` {r Load Dataset}
  # read in the CSV file and make it into a data frame
print(getwd())
  require(data.table)
  system.time(df <- fread("Adult_Arrests.csv"))
  df <- data.frame(df)
df <- subset(df, select= -c(2, 7, 8)) #removing CCN and Arrest Number due to them being fictious 
summary(df)

```

```{r Converting to Factors}
df$Race <- as.factor(df$Race)
df$ETHNICITY <- as.factor(df$ETHNICITY)
df$SEX <-as.factor(df$SEX)
df$CATEGORY <- as.factor(df$CATEGORY)
df$DESCRIPTION <-as.factor(df$DESCRIPTION)
# str(df)
```


```{r Basic Information about Data}
AgeSum <- summary(df$AGE)
YearSum <- summary(df$YEAR)
MonthSum <- summary(df$Month)
Female_Ar <- subset(df, SEX == 3)
Male_Ar <- subset(df, SEX ==2)
Unk_Ar <- subset(df, SEX ==1)
```  

This data set contains information on `r nrow(df)` felony arrests from the District of Columbia between the years 2013-2017. The information provided by this data set is readily available on the District of Columbia’s Police Department website. According to the dossier attached to this dataset, a felony arrest is defined as the police taking into custody a person that is suspected of having committed a crime. Although an arrest occurred, information about the conviction for the crime was not included. As a result, the modeling performed in this report has been done with the assumption that there is a 100% conviction rate for all arrests. This is unlikely to occur so some deviation from the model can be expected. 

In this dataset, the crimes recorded here include boating violations, disorderly conduct, arson, homicide and many more. These crimes range in severity from felony convictions to misdemeanors. Of the crimes committed, `r nrow(Female_Ar)` were the result of the arrest of females, `r nrow(Male_Ar)` were males and `r nrow(Unk_Ar)` of unknown gender orientation. Because there are far more males present in this data set, the model built here may be more applicable to the male population. The mean age of those arrested was 34.8 years old. 

For this report, we have sought to answer the question: **“Is it possible to accurately profile who is likely to commit a crime in DC?”** In the last few years, the use of racial profiling by law enforcement has been a controversial topic. Historically, police officers have had a large presence within primarily African American and Latino communities under the assumption that these areas are plagued with crime. As a result of this, there have many incidences of unlawful arrests and police brutality. 

### Structure of the Data Set 

Within this dataset there was originally 28 variables. ObjectID was removed from this dataset because it corresponded only to the row number. Similarly, CCN and Arrest_Number were removed because these values corresponded to specifics about the arrest that were encrypted due to privacy concerns. After scrubbing the data, the variables left are shown below.

```{r structure, results ="show"}
str(df)
```

```{r Counting missing values, results="show"}
sapply(df, function(x) sum(is.na(x)))
```
Out of the 25 variables present in the cleaned up data set, 12 did not have any missing values. Twelve variables presented up to 2300 missing values. Considering the total number of points in this particular file (153,687), the aformentioned missing values is considerably small. Only one of the variables (Ethnicity) had a large number of missing values (45,899) and, for this reason, was not used in the statistical studies. 


## Correlation of Age and Type of Crime 
```{r Is there any correlation between age and type of crime comitted?, results= "show"}

sex_category <- data.frame(df[c("SEX", "CATEGORY")])
sexcat_Logit <- glm(SEX ~ CATEGORY, data = sex_category, family = "binomial")
summary(sexcat_Logit)
```

By looking at the results from the Logit analysis, we can see that only 3 out of 28 categories are statistically significant. These categories are liquor Law Violations, damage to Property, burglary, offenses against family & children, and arson. As for the three statistically significant variables, category 29 has the lowest p-value suggesting a strong association of the sex of the offender with the probability of comitting arson.


```{r Histogram of age at which crimes were committed, results= "show"}
hist(df$AGE, # histogram
 col="yellow", # column color
 border="black",
 breaks = 25,
 prob = TRUE, # show densities instead of frequencies
 xlim = c(10,80),
 xlab = "Age",
 main = "Age at which crimes are comitted")
lines(density(df$AGE), # density plot
 lwd = 2, # thickness of line
 col = "blue")
```

The histogram and density lines indicate that the age at which the crimes are committed in DC seems to follow a bimodal distribution. Most of the crimes are committed by people with 18 years old. There is a steep decrease on crimes committed by people with ages between 25 and 42 years old. However, this trend is not sustained and the number of crimes  committed by people with ~ 50 years old increases again and then decreases rapidly. 


```{r Can age be used as a predictor for the type of crime committed, results= "show"}
#
df1 <- df[c("AGE", "CATEGORY")]

levels(df1$CATEGORY) <- c("Narcotics", "Release Violations/Fugitive", "Robbery", 
"Prostitution", "Traffic Violations", "Theft", "Driving While Intoxicated", "Property Crimes", "Disorderly Conduct", "Simple Assault", "Other Crimes","Liquor Law Violations", "Damage to Property", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Burglary", "Offenses Against Family & Children", "Vending Violations", "Sex Abuse", "Kidnapping", "Sex Offenses", "Fraud and Financial Crimes", "Weapon Violations", "Theft from Auto", "Homicide", "Motor Vehicle Theft","Gambling","Arson")


ggplot(df1, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=1,show.legend = FALSE,
                outlier.size=1) + ggtitle("Type of Crime Comitted according to age") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
The Box-Whiskers Plot is very useful to understand the data distribution and to vizualize outliers. Here we can see that all but one category (Vending violations) is right-skewed. In a positive-skewed distribution, the mean and median are shifted to higher values when compared to the mode. In the plot above, the right-skewed distribution is most likely caused by the outliers. 

Many assumptions can be made from this Box-Whiskers plot. For instance, robbery, Weapon Violations, homicide and gambling were mostly committed by younger people (late 20's early 30's). An opposite observation can be seen for Vending Violations, which was mostly commited by people in their 40's and 50's. 

### Violent Crimes Committed in Foggy Bottom/Great DC 
```{r Violent crimes comitted at Foggy Botton/DC, results="show"}
loadPkg("mapproj")
loadPkg("ggmap")
#install.packages("mapproj")
#install.packages("ggmap")
#install.packages("DeducerSpatial")
require(maps)
require(ggmap)
register_google(key = "AIzaSyA1A_PfatMAo3MAAL_PcctBHkQKxii0Rto")

violent_crimes <- subset(df, CATEGORY != 2)

# Rank violent crimes
violent_crimes$CATEGORY <- factor(violent_crimes$CATEGORY, levels = c(3,20,26))

# Code numbers: 3- Robbery, 20- Sex abuse, 26- Homicide


dcmapzoom <- qmap('gwu', zoom = 14, color = 'color')

dcmapzoom +geom_point(aes(x = OFFENSE_LONGITUDE, y = OFFENSE_LATITUDE, size = CATEGORY,colour = CATEGORY), data = violent_crimes )+ theme(legend.box = "vertical",legend.position = "topleft")

```

The map above contains the crime committed between 2013-2017. The map is zoomed in Foggy Botton/Dupont Circle/George Town/downtown areas and the red, green, and blue circle corresponds to "Robbery", Sex Abuse, and Homicide respectivally. Foggy Bottom and downtown is well knwon for being relatively safe. This map allows to visualize where the crimes were committed in the investigated area. We can note that robbery is fairly commom in different areas of DC, including GWU premises. However, only few sex abuse and homicides were reported in that area. 

```{r Zoomed out map}

dcmap <- qmap('gwu', zoom = 12, color = 'color')

dcmap +geom_point(aes(x = OFFENSE_LONGITUDE, y = OFFENSE_LATITUDE, size = CATEGORY,colour = CATEGORY), data = violent_crimes )+ theme(legend.box = "vertical",legend.position = "topleft")
```

Sadly, the zoomed out map shows a different trend in DC's neighborhoods. Sex Abuse and homicide crimes increases tremendously in the aformentioned areas. There is an extreme rise in the numbers of homicides and sex abuse in areas like Anacostia, Takoma Park, and Brentwood vicinities. However, the number of sex abuse is notably higher than homicides in the Southwest region. 

### Do people commit crime in their own neighborhood?
```{r Do people commit crime in their neighborhood1}

Liv_OutState <- subset(df, DEFENDANT_PSA == 0) # Arrests of People from out-of-state 
Liv_DC <- subset(df, DEFENDANT_PSA > 100) # Arrests of DC Residents 
Home_Crime <- subset(df, DEFENDANT_PSA == OFFENSE_PSA) #The crime was committed in the area that the defendent lives. 

df2 <- Home_Crime 
levels(df2$CATEGORY) <- c("Narcotics", "Release Violations/Fugitive", "Robbery", 
"Prostitution", "Traffic Violations", "Theft", "Driving While Intoxicated", "Property Crimes", "Disorderly Conduct", "Simple Assault", "Other Crimes","Liquor Law Violations", "Damage to Property", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Burglary", "Offenses Against Family & Children", "Vending Violations", "Sex Abuse", "Kidnapping", "Sex Offenses", "Fraud and Financial Crimes", "Weapon Violations", "Theft from Auto", "Homicide", "Motor Vehicle Theft","Gambling","Arson")

Violet_Crime <- filter(Home_Crime, CATEGORY == 2|3)
levels(Violet_Crime$CATEGORY)
```

To further understand our SMART question, we wanted to examine whether people are more likely to commit a felony offense in their own neighborhood. Based on our findings, `r nrow(Home_Crime)` people, or 24%, of those that committed a crime in DC did so in their own neighborhood, while 76% of crimes were committed by residence of DC in a different neighborhood or by out-of-staters. In order to further understand these findings, we also sought to determine if the crimes committed closer to home were more violent offenses, white collar crimes or misdemeanors. For our purposes, we have defined violent crimes as including Aggravated assault, Assault on a Police Officer, Assault with a Dangerous Weapon, Sex Offenses, Kidnapping, Sex Abuse, Homicide, Weapon violations and Arson.   

```{r Do people commit crime in their neighborhood}
Liv_OutState <- subset(df, DEFENDANT_PSA == 0) # Arrests of People from out-of-state 
Liv_DC <- subset(df, DEFENDANT_PSA > 100) # Arrests of DC Residents 
Home_Crime <- subset(df, DEFENDANT_PSA == OFFENSE_PSA) #DC Resident Commits crime in the area that they live  
Out_Crime_DCres <- subset(Liv_DC, DEFENDANT_PSA!= OFFENSE_PSA) #The crime was committed by a DC resident in an area other than their home. 

Violent_Crime <- Home_Crime %>% filter(CATEGORY == 10 | CATEGORY ==14 | CATEGORY ==15 |CATEGORY == 16 | CATEGORY == 20 | CATEGORY== 22 | CATEGORY == 24 | CATEGORY == 26) %>% droplevels() # Filtering for violent crimes and dropping other levels for DC residents who committed crime in their own neighborhood.

levels(Violent_Crime$CATEGORY) <- c("Simple Assault", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Sex Abuse", "Sex Offenses", "Weapon Violations", "Homicide")

Violent_Crime_Plot <- ggplot(Violent_Crime, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=1,show.legend = FALSE,
                outlier.size=2) + ggtitle("DC Residents who Comitted A Violent Crime Within their Neighborhood")  + theme(axis.text.x = element_text(angle = 30, hjust = 1)) 

Outsider_Crime <- Out_Crime_DCres %>% filter(CATEGORY == 10 | CATEGORY ==14 | CATEGORY ==15 |CATEGORY == 16 | CATEGORY == 20 | CATEGORY== 22 | CATEGORY == 24 | CATEGORY == 26 & DEFENDANT_PSA!=0) %>% droplevels() # Filtering for violent crimes and dropping other levels. 

levels(Outsider_Crime$CATEGORY) <- c("Simple Assault", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Sex Abuse", "Sex Offenses", "Weapon Violations", "Homicide")


Outsider_Crime_Plot <- ggplot(Outsider_Crime, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=1,show.legend = FALSE,
                outlier.size=2) + ggtitle("DC Residents Who comitted a Violent Crime outside their neighborhood")  + theme(axis.text.x = element_text(angle = 30, hjust = 1)) 

OutState <- Liv_OutState %>% filter(CATEGORY == 10 | CATEGORY ==14 | CATEGORY ==15 |CATEGORY == 16 | CATEGORY == 20 | CATEGORY== 22 | CATEGORY == 24 | CATEGORY == 26) %>% droplevels() # Filtering for violent crimes and dropping other levels. 

levels(OutState$CATEGORY) <- c("Simple Assault", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Sex Abuse", "Sex Offenses", "Weapon Violations", "Homicide")

OutState_Plot <- ggplot(OutState, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=1,show.legend = FALSE,
                outlier.size=2) + ggtitle("Out-of-Staters Who comitted a Violent Crime in DC")  + theme(axis.text.x = element_text(angle = 30, hjust = 1)) 

```

``` {r, results = "show"}
Violent_Crime_Plot
Outsider_Crime_Plot
OutState_Plot
```
Interestingly, we found that 24.9% of the arrests made between 2013-2017 were of people who committed said crime within in their own neighborhood. Of this percentage, 51.3% of these crimes were the result of a violent crime. In comparison, 26.5% of the arrests documented here were made by people who lived outside DC or were out-of-staters, while the remaining portion was committed by DC residents who committed a crime in a different neighborhood. Of the crimes that out-of-staters were convicted of only 23.6% were violent. This could indicate that people are more likely to commit a violent crime within their own neighborhood. Similarly, the age spread for the violent crimes committed by those that live within DC in their own neighborhoods tended to be larger than for those who committed a crime in a different neighborhood. 

To better understand the relationship of age and the type of crime committed we have created a density plot that is shown below. Based on this density plot crimes such as narcotics, arson, sex offenses, and gambling are more likely to be committed by people that are younger in age. Felonies involving offenses against family and children, liquor laws, and damage to property occur later in life. If this trend is true, younger people should be profiled more by police as having potentially committed a violent crime.

```{r Does age affect Crime Committed}
ageanova <- aov(AGE ~ CATEGORY, data=df)
ageanovasum <- summary(ageanova)
ageanovasum

ggplot(df1, aes(AGE, fill = CATEGORY)) + geom_density(alpha = 0.2) 
```


## Building a Multivariate Model 

Here, we have sought to determine whether police can accurately determine the type of crime a person is likely to commit based on variables present in this dataset, such as age, race, and ethnicity. First, we have sought to pick the features of this model through using the Bayesian information criterion (BIC). In analyzing the BIC plot shown below, we sought to use the model with the fewest predictors and the lowest BIC score. Based on this, the model that we will build here is hour, age, defendant PSA, race, sex, arrest PSA, arrest district, offense PSA and offense district.

```{r Feature Selection}
loadPkg("leaps")
Features <- subset(df, select= -c(2, 13, 16:19, 22:25)) #removing coordinates of arrest, description of arrest and coordinates of offense. 
colnames(Features) <- c("ID", "Month", "Day", "Hour", "Age", "Defendant_PSA", "Defendant_District", "Race", "Ethnicity", "Sex", "Category", "Arrest_PSA", "Arrest_District", "Offense_PSA", "Offense_District")


Features <- na.omit(Features)

reg.best10 <- regsubsets(Category~. , data = Features, nvmax = 10, nbest = 2, method = "exhaustive")
plot(reg.best10, scale = "bic", main = "BIC")
summary(reg.best10)

```

To develop a multivariate model, we have split the data set into a training and test set. The training data set contained 67% of the values, while the test set contained 33% of the values. Observations were separated randomly, and the data set was scaled to the center.

```{r Creating training Set }
loadPkg("class")
set.seed(1)
Features[] <- lapply(Features, as.numeric) # Converting everything to a numeric 
ScaledFeatures <- as.data.frame(scale(Features[1:15], center = TRUE, scale = TRUE))                

Feature_Sample <- sample(2, nrow(ScaledFeatures), replace=TRUE, prob=c(0.67, 0.33))

Training <- ScaledFeatures[Feature_Sample==1, 1:15]
Test <- ScaledFeatures[Feature_Sample==2, 1:15]


```

The multivariate model was produced using the `lda()` function from the MASS package. Linear discriminate analysis (LDA) seeks to find a linear combination of features that can be used to characterize two or more classes of an event. Essentially this model is attempting to recognize a pattern between the physical variables to predict the crime committed. The coefficients of linear discriminants of this model are displayed below. Each LD can be multiplied by the predictor variable to determine the score for that respondent, which can then be used to compute the posterior probability of class membership. 

```{r Logit, results='show'}
loadPkg("MASS")
logit1 <- lda(Category ~ Hour + Age + Defendant_PSA + Race + Sex + Offense_PSA + Offense_District, data = Training)

Prediction <- logit1 %>% predict(Test)
exp(coef(logit1))

```

## Conclusion 


Linear discriminate analysis assumes that the density of the data is gaussian and that all classes have covariance. This model has been shown to be well suited for multi-class analysis and similar to PCA it can be used as a dimensionality reduction technique. Despite this, when the amount of data for each arrest type is imbalanced in the training set, the model may be unable to accurately classify the observations in the test set. This is an accuracy problem that we will experience when using our model, because there are far less violent crimes in comparison to misdemeanors such as narcotics.  Also, the LDA model requires a defined dimension. Higher-order interactions that may exist between the arrest types may therefore not be captured accurately by this model. 

In conclusion, this data set has been found to have a wide variety of arrest types. Based on the findings in this report, we do not feel that the DC police are able to predict the profile of a person who is more likely to commit a crime in DC. There are two reasons why this is impossible. First, based on the BIC this model needs to include features such as the arrest district, and a person’s one police district. Often when police are profiling a potential criminal, they are making their arrests based on physical descriptions such as race and approximate age. As a result, our model would not actually be able to be employed by officers on the street. Similarly, in agreement with the literature “community policing” often involves some type of bias. To demonstrate the ability of a police officer to predict the crime a person may have committed the model would have to include some type of bias variable. Based on this current data set, community policing should not be used. 
