---
title: "Analysis of Crime within the District of Columbia"
author: "Ashley Frankenfield, Gessica Vasconcelos"
date: "4/3/2020"
output: 
  rmdformats::readthedown:
    html_document:
      number_sections: TRUE
      self_contained: TRUE
      lightbox: TRUE
      gallery: FALSE
      highlight: tango 
      thumbnails: true
      toc_float: TRUE
      use_bookdown: TRUE
      code_folding: hide 
      max-width: 1800px
    pdf_document:
      toc: yes
      toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'hide', message = F, echo=F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
```

```{r basicfcn}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# note that using this function requires quotes around the package name, as you would when installing packages.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
# unload/detact package when done using it 
detach_package = function(pkg, character.only = FALSE) { if(!character.only) { pkg <- deparse(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```


```{r outlierKD2}
outlierKD2 <- function(df, var, rm=FALSE) { 
    #' Original outlierKD functino by By Klodian Dhana,
    #' https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/
    #' Modified to have third argument for removing outliers inwtead of interactive prompt, 
    #' and after removing outlier, original df will not be changed. The function returns the new df, 
    #' which can be saved as original df name if desired.
    #' Check outliers, and option to remove them, save as a new dataframe. 
    #' @param df The dataframe.
    #' @param var The variable in the dataframe to be checked for outliers
    #' @param rm Boolean. Whether to remove outliers or not.
    #' @return The dataframe with outliers replaced by NA if rm==TRUE, or df if nothing changed
    #' @examples
    #' outlierKD2(mydf, height, FALSE)
    #' mydf = outlierKD2(mydf, height, TRUE)
    #' mydfnew = outlierKD2(mydf, height, TRUE)
    dt = df # duplicate the dataframe for potential alteration
    var_name <- eval(substitute(var),eval(dt))
    na1 <- sum(is.na(var_name))
    m1 <- mean(var_name, na.rm = T)
    par(mfrow=c(2, 2), oma=c(0,0,3,0))
    boxplot(var_name, main="With outliers")
    hist(var_name, main="With outliers", xlab=NA, ylab=NA)
    outlier <- boxplot.stats(var_name)$out
    mo <- mean(outlier)
    var_name <- ifelse(var_name %in% outlier, NA, var_name)
    boxplot(var_name, main="Without outliers")
    hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
    title("Outlier Check", outer=TRUE)
    na2 <- sum(is.na(var_name))
    cat("Outliers identified:", na2 - na1, "\n")
    cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "\n")
    cat("Mean of the outliers:", round(mo, 2), "\n")
    m2 <- mean(var_name, na.rm = T)
    cat("Mean without removing outliers:", round(m1, 2), "\n")
    cat("Mean if we remove outliers:", round(m2, 2), "\n")
    
    # response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
    # if(response == "y" | response == "yes"){
    if(rm){
        dt[as.character(substitute(var))] <- invisible(var_name)
        #assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
        cat("Outliers successfully removed", "\n")
        return(invisible(dt))
    } else {
        cat("Nothing changed", "\n")
        return(invisible(df))
    }
}
```

```{r loadPkg}
  loadPkg("corrplot")
  loadPkg("dplyr")
  loadPkg("ggplot2")
  loadPkg("tidyverse")
  loadPkg("sqldf")
  loadPkg('tidyr')
```

## Introduction 
``` {r Load Dataset}
  # read in the CSV file and make it into a data frame
print(getwd())
  require(data.table)
  system.time(df <- fread("Adult_Arrests.csv"))
  df <- data.frame(df)
df <- subset(df, select= -c(2, 7, 8)) #removing CCN and Arrest Number due to them being fictious 
summary(df)

```

```{r Converting to Factors}
df$Race <- as.factor(df$Race)
df$ETHNICITY <- as.factor(df$ETHNICITY)
df$SEX <-as.factor(df$SEX)
df$CATEGORY <- as.factor(df$CATEGORY)
df$DESCRIPTION <-as.factor(df$DESCRIPTION)
# str(df)
```

```{r Counting missing values and removing not relevant variables}
sapply(df, function(x) sum(is.na(x)))
```

```{r Basic Information about Data}
AgeSum <- summary(df$AGE)
YearSum <- summary(df$YEAR)
MonthSum <- summary(df$Month)
Female_Ar <- subset(df, SEX == 3)
Male_Ar <- subset(df, SEX ==2)
Unk_Ar <- subset(df, SEX ==1)
```  

This data set contains information on `r nrow(df)` felony arrests from the District of Columbia between the years 2013-2017. The information provided by this data set is readily available on the District of Columbia’s Police Department website. According to the dossier attached to this dataset, a felony arrest is defined as the police taking into custody a person that is suspected of having committed a crime. Although an arrest occurred, information about the conviction for the crime was not included. As a result, the modeling performed in this report has been done with the assumption that there is a 100% conviction rate for all arrests. This is unlikely to occur so some deviation from the model can be expected. 

In this dataset, the crimes recorded here include boating violations, disorderly conduct, arson, homicide and many more. These crimes range in severity from felony convictions to misdemeanors. Of the crimes committed, `r nrow(Female_Ar)` were the result of the arrest of females, `r nrow(Male_Ar)` were males and `r nrow(Unk_Ar)` of unknown gender orientation. Because there are far more males present in this data set, the model built here may be more applicable to the male population. The mean age of those arrested was 34.8 years old. 

For this report, we have sought to answer the question: **“Is it possible to accurately profile who is likely to commit a crime in DC?”** In the last few years, the use of racial profiling by law enforcement has been a controversial topic. Historically, police officers have had a large presence within primarily African American and Latino communities under the assumption that these areas are plagued with crime. As a result of this, there have many incidences of unlawful arrests and police brutality. 

### Structure of the Data Set 

Within this dataset there was originally 28 variables. ObjectID was removed from this dataset because it corresponded only to the row number. Similarly, CCN and Arrest_Number were removed because these values corresponded to specifics about the arrest that were encrypted due to privacy concerns. After scrubbing the data, the variables left are shown below.

```{r structure, results ="show"}
str(df)
```

## Correlation of Age and Type of Crime 
```{r Is there any correlation between age and type of crime comitted?}

sex_category <- data.frame(df[c("SEX", "CATEGORY")])
sexcat_Logit <- glm(SEX ~ CATEGORY, data = sex_category, family = "binomial")
summary(sexcat_Logit)
```

By looking at the results from the Logit analysis, we can see that only 5 out of 28 categories are statistically significant. These categories are liquor Law Violations, damage to Property, burglary, offenses against family & children, and arson. As for thee statistically significant variables, category 29 has the lowest p-value suggesting a strong association of the sex of the offender with the probability of comitting arson. The negative coefficient for this predictor suggests that all other variables being equal, the gender is less likely to male passenger is less likely to have survived.


```{r}
hist(df$AGE, # histogram
 col="yellow", # column color
 border="black",
 breaks = 25,
 prob = TRUE, # show densities instead of frequencies
 xlim = c(10,80),
 xlab = "Age",
 main = "Age at which crimes are comitted")
lines(density(df$AGE), # density plot
 lwd = 2, # thickness of line
 col = "blue")
```

The histogram and density lines indicate that the age at which the crimes are committed in DC seems to follow a bimodal distribution. Most of the crimes are committed by people with 18 years old. There is a steep decrease on crimes committed by people with aged between 25 and 42 years old. However, this trend is not sustained and the number crimes increases committed by people with ~ 50 years old increases again. 

```{r Can age be used as a predictor for the type of crime commited}

df1 <- df[c("AGE", "CATEGORY")]

levels(df1$CATEGORY) <- c("Narcotics", "Release Violations/Fugitive", "Robbery", 
"Prostitution", "Traffic Violations", "Theft", "Driving While Intoxicated", "Property Crimes", "Disorderly Conduct", "Simple Assault", "Other Crimes","Liquor Law Violations", "Damage to Property", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Burglary", "Offenses Against Family & Children", "Vending Violations", "Sex Abuse", "Kidnapping", "Sex Offenses", "Fraud and Financial Crimes", "Weapon Violations", "Theft from Auto", "Homicide", "Motor Vehicle Theft","Gambling","Arson")


ggplot(df1, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=1,show.legend = FALSE,
                outlier.size=1) + ggtitle("Type of Crime Comitted according to age") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## Violent Crime Commited in Foggy Bottom 
```{r Violent crimes comitted at Foggy Botton}
loadPkg("mapproj")
loadPkg("ggmap")
#install.packages("mapproj")
#install.packages("ggmap")
#install.packages("DeducerSpatial")
require(maps)
require(ggmap)
#register_google(key = "AIzaSyA1A_PfatMAo3MAAL_PcctBHkQKxii0Rto")

violent_crimes <- subset(df, CATEGORY != 2)

# rank violent crimes
violent_crimes$CATEGORY <- factor(violent_crimes$CATEGORY, levels = c(16,20,26))
# 3- robbery, 14- aggravated assault, 20- Sex abuse, 26- homicide, 1-Narcotics , 12-Liquor Law Violations, 28-gambling


#dcmap <- qmap('gwu', zoom = 12, color = 'color')

#dcmap +geom_point(aes(x = OFFENSE_LONGITUDE, y = OFFENSE_LATITUDE, size = CATEGORY,colour = CATEGORY), data = violent_crimes )+ theme(legend.box = "vertical",legend.position = "topleft")

```

```{r Is there a relationship between age and the type of crime comitted? }

fit <- lm(AGE ~ CATEGORY, data = df)
summary(fit)
```

## Do people commit crimes in their own neighborhoods? 

```{r Do people commit crime in their neighborhood1}
Liv_OutState <- subset(df, DEFENDANT_PSA == 0) # Arrests of People from out-of-state 
Liv_DC <- subset(df, DEFENDANT_PSA > 100) # Arrests of DC Residents 
Home_Crime <- subset(df, DEFENDANT_PSA == OFFENSE_PSA) #The crime was committed in the area that the defendent lives. 

df2 <- Home_Crime 
levels(df2$CATEGORY) <- c("Narcotics", "Release Violations/Fugitive", "Robbery", 
"Prostitution", "Traffic Violations", "Theft", "Driving While Intoxicated", "Property Crimes", "Disorderly Conduct", "Simple Assault", "Other Crimes","Liquor Law Violations", "Damage to Property", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Burglary", "Offenses Against Family & Children", "Vending Violations", "Sex Abuse", "Kidnapping", "Sex Offenses", "Fraud and Financial Crimes", "Weapon Violations", "Theft from Auto", "Homicide", "Motor Vehicle Theft","Gambling","Arson")

Violet_Crime <- filter(Home_Crime, CATEGORY == 2|3)
levels(Violet_Crime$CATEGORY)
```

To further understand our SMART question, we wanted to examine whether people are more likely to commit a felony offense in their own neighborhood. Based on our findings, `r nrow(Home_Crime)` people, or 24%, of those that committed a crime in DC did so in their own neighborhood, while 76% of crimes were committed by residence of DC in a different neighborhood or by out-of-staters. In order to further understand these findings, we also sought to determine if the crimes committed closer to home were more violent offenses, white collar crimes or misdemeanors. For our purposes, we have defined violent crimes as including Aggravated assault, Assault on a Police Officer, Assault with a Dangerous Weapon, Sex Offenses, Kidnapping, Sex Abuse, Homicde, Weapon violations and Arson.  


```{r Feature Selection}
loadPkg("leaps")
Features <- subset(df, select= -c(2, 13, 16:19, 22:25)) #removing coordinates of arrest, description of arrest and coordinates of offense. 
colnames(Features) <- c("ID", "Month", "Day", "Hour", "Age", "Defendant_PSA", "Defendant_District", "Race", "Ethnicity", "Sex", "Category", "Arrest_PSA", "Arrest_District", "Offense_PSA", "Offense_District")


Features <- na.omit(Features)

reg.best10 <- regsubsets(Category~. , data = Features, nvmax = 10, nbest = 2, method = "exhaustive")
plot(reg.best10, scale = "bic", main = "BIC")
summary(reg.best10)

```

```{r Do people commit crime in their neighborhood}
Liv_OutState <- subset(df, DEFENDANT_PSA == 0) # Arrests of People from out-of-state 
Liv_DC <- subset(df, DEFENDANT_PSA > 100) # Arrests of DC Residents 
Home_Crime <- subset(df, DEFENDANT_PSA == OFFENSE_PSA) #DC Resident Commits crime in the area that they live  
Out_Crime_DCres <- subset(Liv_DC, DEFENDANT_PSA!= OFFENSE_PSA) #The crime was committed by a DC resident in an area other than their home. 

Violent_Crime <- Home_Crime %>% filter(CATEGORY == 10 | CATEGORY ==14 | CATEGORY ==15 |CATEGORY == 16 | CATEGORY == 20 | CATEGORY== 22 | CATEGORY == 24 | CATEGORY == 26) %>% droplevels() # Filtering for violent crimes and dropping other levels for DC residents who committed crime in their own neighborhood.

levels(Violent_Crime$CATEGORY) <- c("Simple Assault", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Sex Abuse", "Sex Offenses", "Weapon Violations", "Homicide")

Violent_Crime_Plot <- ggplot(Violent_Crime, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=1,show.legend = FALSE,
                outlier.size=2) + ggtitle("DC Residents who Comitted A Violent Crime Within their Neighborhood")  + theme(axis.text.x = element_text(angle = 30, hjust = 1)) 

Outsider_Crime <- Out_Crime_DCres %>% filter(CATEGORY == 10 | CATEGORY ==14 | CATEGORY ==15 |CATEGORY == 16 | CATEGORY == 20 | CATEGORY== 22 | CATEGORY == 24 | CATEGORY == 26 & DEFENDANT_PSA!=0) %>% droplevels() # Filtering for violent crimes and dropping other levels. 

levels(Outsider_Crime$CATEGORY) <- c("Simple Assault", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Sex Abuse", "Sex Offenses", "Weapon Violations", "Homicide")


Outsider_Crime_Plot <- ggplot(Outsider_Crime, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=1,show.legend = FALSE,
                outlier.size=2) + ggtitle("DC Residents Who comitted a Violent Crime outside their neighborhood")  + theme(axis.text.x = element_text(angle = 30, hjust = 1)) 

OutState <- Liv_OutState %>% filter(CATEGORY == 10 | CATEGORY ==14 | CATEGORY ==15 |CATEGORY == 16 | CATEGORY == 20 | CATEGORY== 22 | CATEGORY == 24 | CATEGORY == 26) %>% droplevels() # Filtering for violent crimes and dropping other levels. 

levels(OutState$CATEGORY) <- c("Simple Assault", "Aggravated Assault", "Assault on a Police Officer", "Assault with a Dangerous Weapon", "Sex Abuse", "Sex Offenses", "Weapon Violations", "Homicide")

OutState_Plot <- ggplot(OutState, aes(x=CATEGORY, y=AGE, fill= CATEGORY)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=1,show.legend = FALSE,
                outlier.size=2) + ggtitle("Out-of-Staters Who comitted a Violent Crime in DC")  + theme(axis.text.x = element_text(angle = 30, hjust = 1)) 

```

We wanted to examine whether or not people are more likely to commit a felony offense in their own neighborhood. Based on our findings, `r nrow(Home_Crime)` people, or 24%, of those that committed a crime in DC did so in their own neighborhood. We wanted to extend these findings and see what crimes people are more likely to commit closer to home. For our purposes, we have distinguished between violent offenses, white collar crimes and misdemenors. 

```{r Counts}
nrow(Home_Crime)
nrow(Liv_OutState)
nrow(Out_Crime_DCres)

#Number of Violent Crimes committed per category 
nrow(Violent_Crime)
nrow(OutState)
nrow(Outsider_Crime)
```

```{r Does age affect Crime Committed}
ageanova <- aov(AGE ~ CATEGORY, data=df)
ageanovasum <- summary(ageanova)
ageanovasum

ggplot(df1, aes(AGE, fill = CATEGORY)) +  geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity')
ggplot(df1, aes(AGE, fill = CATEGORY)) + geom_density(alpha = 0.2) 
```
      
```{r does sex affect Crime type, results='show' }
loadPkg("gplots")
gendercontable = xtabs( ~ CATEGORY + SEX, data = df)
chisqsex = chisq.test(gendercontable)
chisqsex
gendercontable
```

## Building a Multinomial Model 

```{r Creating training Set }
loadPkg("class")
set.seed(1)
Features[] <- lapply(Features, as.numeric) # Converting everything to a numeric 
ScaledFeatures <- as.data.frame(scale(Features[1:15], center = TRUE, scale = TRUE))                

Feature_Sample <- sample(2, nrow(ScaledFeatures), replace=TRUE, prob=c(0.67, 0.33))

Training <- ScaledFeatures[Feature_Sample==1, 1:15]
Test <- ScaledFeatures[Feature_Sample==2, 1:15]


Feature_Training <- Features[Feature_Sample==1, 11]
Feature_Test <- Features[Feature_Sample==1, 11]

#prc_test_pred <- knn(train = Training, test = Test,cl = Feature_Training, k=7)
```

```{r Logit, results='show'}
loadPkg("MASS")
logit1 <- lda(Category ~ Hour + Age + Defendant_PSA + Race + Sex + Offense_PSA + Offense_District, data = Training)

Prediction <- logit1 %>% predict(Test)
logit1
exp(coef(logit1))

```




