---
title: "Project 2"
author: "Gessica Vasconcelos"
date: "4/3/2020"
output: 
  rmdformats::readthedown:
    html_document:
      number_sections: TRUE
      self_contained: TRUE
      lightbox: TRUE
      gallery: FALSE
      highlight: tango 
      thumbnails: true
      toc_float: TRUE
      use_bookdown: TRUE
      code_folding: hide 
      max-width: 1800px
    pdf_document:
      toc: yes
      toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'hide', message = F, echo=F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
```

```{r basicfcn}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# note that using this function requires quotes around the package name, as you would when installing packages.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }
# unload/detact package when done using it 
detach_package = function(pkg, character.only = FALSE) { if(!character.only) { pkg <- deparse(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r loadPkg}
  loadPkg("corrplot")
  loadPkg("dplyr")
  loadPkg("ggplot2")
  loadPkg("tidyverse")
  loadPkg("sqldf")
  loadPkg('tidyr')
```

``` {r Load Dataset}
  # read in the CSV file and make it into a data frame
print(getwd())
  require(data.table)
  system.time(df <- fread("Adult_Arrests.csv"))
  df <- data.frame(df)
str(df)
```

```{r Converting to Factors}

#I believe I'm using the same file as you as you (the one you sent me on the email) but I was getting an error so I just changed the names to match the ones in the file.

df$Race <- as.factor(df$Race)
df$ETHNICITY <- as.factor(df$ETHNICITY)
df$SEX <-as.factor(df$SEX)
df$CATEGORY <- as.factor(df$CATEGORY)
df$DESCRIPTION <-as.factor(df$DESCRIPTION)
str(df)
```


```{r Counting missing values and removing not relevant variables}
sapply(df, function(x) sum(is.na(x)))
select (df,-c(CCN,ARREST_NUMBER))
```

```{r Basic Information about Data}
AgeSum <- summary(df$AGE)
YearSum <- summary(df$YEAR)
MonthSum <- summary(df$Month)
Female_Ar <- subset(df, SEX == 3)
Male_Ar <- subset(df, SEX ==2)
Unk_Ar <- subset(df, SEX ==1)

```  

This data set contains information on `r nrow(df)` felony arrests from the District of Columbia between the years 2013-2017. A felony arrest involves the police taking into custody a person that is suspected of having committed a crime. Crimes that are less severe often result in misdemeanor charges, while more serious acts can result in a felony charge. In this data set, both type of charges are included. The crimes recorded here include boating violations, disorderly conduct, arson, homicide and many more. These crimes have resulted in the arrest of `r nrow(Female_Ar)` females, `r nrow(Male_Ar)` were males and `r nrow(Unk_Ar)` of unknown gender orientation. The mean age of those arrested is 34.8 years old. 

```{r Do people commit crime in their neighborhood?}
Y2013 <- subset(df, YEAR == 2013)
Y2014 <- subset(df, YEAR == 2014)
Y2015 <- subset(df, YEAR == 2015)
Y2016 <- subset(df, YEAR == 2016)
Y2017 <- subset(df, YEAR == 2017)
Liv_OutState <- subset(df, DEFENDANT_PSA == 0) # Arrests of People from out-of-state 
Liv_DC <- subset(df, DEFENDANT_PSA > 100) # Arrests of DC Residents 
```


```{r Is there any correlation between age and type of crime comitted?}

sex_category <- data.frame(df[c("SEX", "CATEGORY")])
sexcat_Logit <- glm(SEX ~ CATEGORY, data = sex_category, family = "binomial")
summary(sexcat_Logit)
anova(model, test="Chisq")
```

By lookinG at the results from the Logit analysis, we can see that only 5 out of 28 categories are statistically significant. These categories are liquor Law Violations, damage to Property, burglary, offenses against family & children, and arson. As for thee statistically significant variables, category 29 has the lowest p-value suggesting a strong association of the sex of the offender with the probability of comitting arson. The negative coefficient for this predictor suggests that all other variables being equal, the gender is less likely to male passenger is less likely to have survived.


```{r Does the month aff}
#Checking normality of the variable Age
ggplot(data=df, aes(x=AGE)) + 
    geom_histogram( binwidth=2, fill="blue", color="#e9ecef") +
    ggtitle("Distribution of Age")


```
The histogram indicates that the age 

month_cat <- data.frame(df[c("CATEGORY", "AGE")])
monthcat_Logit <- glm(AGE ~ CATEGORY, data = month_cat, family = "binomial")
summary(monthcat_Logit)
anova(monthcat_Logit, test="Chisq")


```

